\forestclaw: Hybrid forest-of-octrees AMR for hyperbolic conservation laws

### 2013-10-31 docs/parco13/paper.tex
Tweaks and new weak scaling table

We provide two examples that demonstrate the \forestclaw code.
The numerical results to date have been designed to verify that
interface layer between \pforest and \forestclaw is sufficiently
flexible and robust.  Of particular importance was ensuring that all
ghost cell transfers (including averaging and
interpolation) are implemented correctly.
The basic \clawpack algorithm and corresponding code are
thoroughly tested and need no further verification in our context.

@@

In \Fig{ompscale}, we show multi-node calculations using 16 OpenMP
threads per node executed by the
%and in future
%iterations on accelerators 
\manyclaw set of patch-based PDE solvers.
%with
%\forestclaw.  
%\manyclaw provides a generic set of interators for Riemann
%problems and related algorithms.

@@

We have examined weak scalability as well by using four times as many MPI
processes
for each level increase, from level 6 at 64 cores to level 11 at 16,384.  Up to
4,096 cores, the exchange times stay consistently below 16\% of the time
integration and the time in seconds per time integration step varies between
.0578 and .0602, which yields a parallel efficiency of 96\%.  For the
data point at 16,384 cores, the efficiency is still 82\%.

\begin{table}
  \begin{center}
    \begin{tabular}{cccc|rr}
      \multicolumn{4}{c|}{strategy} & \multicolumn{2}{c}{wall clock time} \\
      mesh & remesh & partition & time step & $P = 16$ & $P = 256$ \\
      \hline
      uniform & none       & by count  & global   & 3961.  & 256.  \\
      \hline
          AMR & every step & by count  & global   &  252.  &  54.6 \\
          AMR & every 4    & by count  & global   &  178.  &  39.7 \\
      \hline
          AMR & every step & by count  & subcycle &   99.9 &  17.3 \\
          AMR & every 4    & by count  & subcycle &   87.2 &  14.0 \\
      \hline
          AMR & every step & by weight & subcycle &   95.7 &  18.2 \\
          AMR & every 4    & by weight & subcycle &   84.4 & 14.2 \\
      \hline
    \end{tabular}
  \end{center}
\caption{Comparison of different meshing strategies for the advection example
  with a fixed maximum level 8, run on 16 and 256 cores, respectively.  The
  wall clock time includes the whole run of the program beginning to end (file
  I/O disabled).  The adaptive runs converge to a minimum level of 3.}
\label{tab:swirlwall}
\end{table}
Next, we examine how we can reduce the overall wall clock time of a simulation
by switching from uniformly refined meshes to adaptive refinement with the same maximum level,
effectively coarsening where high resolution is not needed.  For an example
with a sharp front as depicted in \Fig{results2d} we expect considerable
savings by AMR, which we confirm in \Tab{swirlwall}.  The wall clock times can
be reduced by up to a factor of 50 by combining AMR with subcycling.  The
adaptive runs are still  load balanced overall, which we infer from
the linear scaling of the time integration alone (not shown) and from the fact
that the weighted partition does not influence the run times.  There is the
tradeoff however, at least in the current implementation, that the adaptive
runs exhibit reduced scalability associated with the ghost patch exchanges
described earlier.  Presently, ghost patch exchanges are made across all levels, even
if only two levels are involved in the exchange.  One enhancement of the code that is
currently being developed is to rewrite the time stepping loop with the
objective of minimizing the parallel neighbor exchanges.

>>>>

We provide two examples that exercise the \forestclaw code.
The numerical results to date have been designed to verify that
interface layer between \pforest and \forestclaw is sufficiently
flexible and robust.  Of particular importance was ensuring that all
ghost cell transfers (including averaging and
interpolation) are implemented correctly.
The basic \clawpack algorithms and the corresponding code are
thoroughly tested and need no further verification in our context.

@@

In \Fig{ompscale}, we show multi-node calculations using 16 OpenMP
threads per node executed by the
%and in future
%iterations on accelerators 
\manyclaw set of patch-based PDE solvers \cite{manyclaw}.
%with
%\forestclaw.  
%\manyclaw provides a generic set of interators for Riemann
%problems and related algorithms.

@@

\begin{table}
  \begin{center}
   \begin{tabular}{l|rrrrrr}
      $P$ (\# total MPI ranks) & 16 & 64 & 256 & 1,024 & 4,096 & 16,384 \\
      \hline
      par.\ efficiency (\%) & 100 & 96 & 96 & 103 & 103 & 82 \\
    \end{tabular}
  \end{center}
  \caption{Weak scalability on a uniform mesh from level 6 at 64 cores to level
    11 at 16,384
    % using
    (16 MPI ranks/node).}%
  \label{tab:weakscale}
\end{table}%
%
We have examined weak scalability as well by using four times as many MPI
processes for each level increase.
% , from level 6 at 64 cores to level 11 at 16,384.
Up to 4,096 cores, the exchange times stay consistently below 16\% of the time
integration,
% and the time in seconds per time integration step varies between
% .0578 and .0602,
and we obtain optimal parallel efficiency.  For
% the data point at
16,384 cores, the efficiency reduces to 82\%
with the current code (see \Tab{weakscale}).

\begin{table}
  \begin{center}
    \begin{tabular}{cccc|rr}
      \multicolumn{4}{c|}{strategy} & \multicolumn{2}{c}{wall clock time} \\
      mesh & remesh & partition & time step & $P = 16$ & $P = 256$ \\
      \hline
      uniform & none       & by count  & global   & 3961.  & 256.  \\
      \hline
          AMR & every step & by count  & global   &  252.  &  54.6 \\
          AMR & every 4    & by count  & global   &  178.  &  39.7 \\
      \hline
          AMR & every step & by count  & subcycle &   99.9 &  17.3 \\
          AMR & every 4    & by count  & subcycle &   87.2 &  14.0 \\
      \hline
          AMR & every step & by weight & subcycle &   95.7 &  18.2 \\
          AMR & every 4    & by weight & subcycle &   84.4 & 14.2 \\
      \hline
    \end{tabular}
  \end{center}
\caption{Comparison of different meshing strategies for the advection example
  with a fixed maximum level 8, run on 16 and 256 cores, respectively.  The
  wall clock time includes the whole run of the program beginning to end (file
  I/O disabled).  The adaptive runs converge to a minimum level of 3.}
\label{tab:swirlwall}
\end{table}
Next, we examine how we can reduce the overall wall clock time of a simulation
by switching from uniformly refined meshes to adaptive refinement with the same maximum level,
effectively coarsening where high resolution is not needed.  For an example
with a sharp front as depicted in \Fig{results2d} we expect considerable
savings by AMR, which we confirm in \Tab{swirlwall}.  The wall clock times can
be reduced by up to a factor of 50 by combining AMR with subcycling.  The
adaptive runs are still  load balanced overall, which we infer from
the linear scaling of the time integration alone (not shown) and from the fact
that the weighted partition does not influence the run times.  There is the
tradeoff however, at least in the current implementation, that the adaptive
runs exhibit reduced scalability associated with the ghost patch exchanges
described earlier.
% One enhancement of the code that is currently being developed is to rewrite
% the time stepping loop with the objective of minimizing the parallel neighbor
% exchanges.
%
% Presently, ghost patch exchanges are made across all levels, even
% if only two levels are involved in the exchange.  One enhancement of the code that is
% currently being developed is to rewrite the time stepping loop with the
% objective of minimizing the parallel neighbor exchanges.
